---
title: "A3: SPL Library Checkouts in relation to the novel, Normal People"
author: "Connor Chan"
output: html_document
---


### Introduction

For Assignment 3, I am examining data from the Seattle Public Library, more specifically checkout data from the novel, Normal People, written by Sally Rooney. The dataset I chose from the public library includes both univariate and multivariate categories including Usage Class, Checkout Type, Material Type, Checkout Year, Checkout Month, number of checkouts, Title, ISBN, author, genre, Publisher, and Publication Year.

I am choosing to examine this dataset because I am currently reading Normal People, and I am interested in learning about the popularity of the novel between its publish date and today.

### Summary Information



```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE, message = FALSE)


# read the CSV file into a variable called "normal_data"
normal_data <- read.csv("file:///Users/connorchan/Library/Messages/Attachments/dd/13/00961DEA-54E7-47C6-84E3-697CD2A044EA/Normal_People_A3_Checkouts_by_Title.csv")

library("dplyr")
library("stringr")
library("ggplot2")
library("scales")
library(tidyverse)


```

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

For research questions, I drew questions from the inspiration section of the dataset as well as from my personal interests. All of my questions are related to the dataset chosen regarding the novel, Normal People.

1. What is the average number of checkouts?

Answer: 122.8106


2. What month has the most checkouts?

Answer: May; 1644


3. What year has the least amount of checkouts?

Answer: 2023; 151 checkouts


4. What is the most common checkout material type?

Answer: Audiobooks

5. What is the least common material type?

Answer: Books


These 5 questions will allow me to properly examine the data, allowing me to make new discoveries about checkout data regarding the novel, Normal People at the Seattle Public Library.

```{r, echo = TRUE, p1}

```

### The Dataset

#### Who collected/published the data?
The Seattle Public Library

#### What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?


#### How was the data collected or generated?
The data was collected and generated by the Seattle Public Library database

#### Why was the data collected?
The data was collected to look at the checkout data at the Seattle Public Library, to allow open knowledge access to the public, this data allows people to visualize trends in data and various categories.

#### What, if any, ethical questions do you need to consider when working with this data?
The data collected from the Seattle Public Library between April 2019 and February 2023 is unbiased data, collected through checkout machines. However, it is important to consider possible ethical questions that may change data accuracy, some of which include: people taking books without proper checking out and machine malfunction. When working with this data to answer the questions at hand, it is important to weigh the accuracy of data. In this specific instance, it is most likely that the data is accurate.

#### What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning = FALSE, code = readLines("chart1_example.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, code = readLines("chart2_example.R")}
```


