---
title: "A3: SPL Library Checkouts in relation to the novel, Normal People"
author: "Connor Chan"
output: html_document
---


### Introduction

For Assignment 3, I am examining data from the Seattle Public Library, more specifically checkout data from the novel, Normal People, written by Sally Rooney. The dataset I chose from the public library includes both univariate and multivariate categories including Usage Class, Checkout Type, Material Type, Checkout Year, Checkout Month, number of checkouts, Title, ISBN, author, genre, Publisher, and Publication Year.

I am choosing to examine this dataset because I am currently reading Normal People, and I am interested in learning about the popularity of the novel between its publish date and today.

### Summary Information



```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# read the CSV file into a variable called "normal_data"
normal_data <- read.csv("file:///Users/connorchan/Library/Messages/Attachments/dd/13/00961DEA-54E7-47C6-84E3-697CD2A044EA/Normal_People_A3_Checkouts_by_Title.csv")

library("dplyr")
library("stringr")
library("ggplot2")
library("scales")
library(tidyverse)

```

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

For research questions, I drew questions from the inspiration section of the dataset as well as from my personal interests. All of my questions are related to the dataset chosen regarding the novel, Normal People.

1. What is the average number of checkouts?

Answer: 122.8106


2. What month has the most checkouts?

Answer: May; 1644 checkouts


3. What year has the least amount of checkouts?

Answer: 2023; 151 checkouts


4. What is the most common checkout material type?

Answer: Audiobooks

5. What is the least common material type?

Answer: Books


These 5 questions will allow me to properly examine the data, allowing me to make new discoveries about checkout data regarding the novel, Normal People at the Seattle Public Library.

```{r, echo = TRUE, p1}

```

### The Dataset

#### Who collected/published the data?
The Seattle Public Library

#### What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
The parameters within the dataset include Usage Class, Checkout Type, Material Type, Checkout Year, Checkout Month, number of checkouts, Title, ISBN, author, genre, Publisher, and Publication Year.

#### How was the data collected or generated?
The data was collected and generated by the Seattle Public Library database

#### Why was the data collected?
The data was collected to look at the checkout data at the Seattle Public Library, to allow open knowledge access to the public, this data allows people to visualize trends in data and various categories.

#### What, if any, ethical questions do you need to consider when working with this data?
The data collected from the Seattle Public Library between April 2019 and February 2023 is unbiased data, collected through checkout machines. However, it is important to consider possible ethical questions that may change data accuracy, some of which include: people taking books without proper checking out and machine malfunction. When working with this data to answer the questions at hand, it is important to weigh the accuracy of data. In this specific instance, it is most likely that the data is accurate.

#### What are possible limitations or problems with this data?   (at least 200 words)
The Normal People dataset from the Seattle Public Library represents a single library. As a result, the data does not accurately reflect checkout patterns across global libraries and data systems. For example, if the library serves a unique demographic or has a specific collection, the observed patterns will not represent all checkout patterns.

Another potential concern with the Normal People dataset is its limited scope - it only covers three years of data. This makes it challenging to analyze trends over time or compare the data to previous years. In my data science class last quarter, I learned that data is the most accurate with as much collected information as possible. As a result, the trends in the Normal People dataset will be most accurate as checkout data continues to be collected.

The Normal People dataset may also suffer from potential errors and inconsistencies in the data collection process. For example, there may be discrepancies in how checkout data has been recorded or there may be different methods used to count checkouts. These manual and technical issues could lead to inaccuracies in the dataset and may impact the validity of conclusions drawn from the data. As a result, it is important to carefully review the data and consider potential sources of error before drawing any definitive conclusions.

### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning=FALSE, message=FALSE}

# Load necessary libraries
library(dplyr)
library(ggplot2)

# Read in the data
normal_data <- read.csv("file:///Users/connorchan/Library/Messages/Attachments/dd/13/00961DEA-54E7-47C6-84E3-697CD2A044EA/Normal_People_A3_Checkouts_by_Title.csv")

# Clean up variable names
colnames(normal_data) <- c("Usage_Class", "Checkout_Type", "Material_Type", "Checkout_Year", "Checkout_Month", "Checkout_Number")

# Convert Checkout_Year and Checkout_Month to a single date variable
normal_data$Date <- as.Date(paste0(normal_data$Checkout_Year, "-", normal_data$Checkout_Month, "-01"))

# Filter for Books, Ebooks, and Audiobooks only
normal_data <- subset(normal_data, Material_Type %in% c("BOOK", "EBOOK", "AUDIOBOOK"))

# Group by Material_Type and Date, and summarize by summing Checkout_Number
normal_data <- aggregate(Checkout_Number ~ Material_Type + Date, data = normal_data, FUN = sum)

# Plot the data using ggplot2
ggplot(normal_data, aes(x = Date, y = Checkout_Number, color = Material_Type)) +
  geom_line() +
  labs(x = "Checkout Date", y = "Number of Checkouts", color = "Material Type") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 year") +
  theme_minimal()


```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning=FALSE, message=FALSE}

# Load necessary libraries
library(dplyr)
library(ggplot2)

# Read in the data
normal_data <- read.csv("file:///Users/connorchan/Library/Messages/Attachments/dd/13/00961DEA-54E7-47C6-84E3-697CD2A044EA/Normal_People_A3_Checkouts_by_Title.csv")

# Clean up variable names
colnames(normal_data) <- c("Usage_Class", "Checkout_Type", "Material_Type", "Checkout_Year", "Checkout_Month", "Checkout_Number")

# Filter for Books, Ebooks, and Audiobooks only
normal_data <- subset(normal_data, Material_Type %in% c("BOOK", "EBOOK", "AUDIOBOOK"))

# Group by Material_Type and Checkout_Year, and summarize by summing Checkout_Number
normal_data <- aggregate(Checkout_Number ~ Material_Type + Checkout_Year, data = normal_data, FUN = sum)

# Plot the data using ggplot2
ggplot(normal_data, aes(x = Checkout_Year, y = Checkout_Number, fill = Material_Type)) +
  geom_bar(stat = "identity") +
  labs(x = "Checkout Year", y = "Number of Checkouts", fill = "Material Type") +
  theme_minimal()

```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)

# Load data
normal_data <- read.csv("file:///Users/connorchan/Library/Messages/Attachments/dd/13/00961DEA-54E7-47C6-84E3-697CD2A044EA/Normal_People_A3_Checkouts_by_Title.csv")

# Filter data for years 2019-2023 and calculate total checkouts by material type
material_counts <- normal_data %>%
  filter(CheckoutYear >= 2019 & CheckoutYear <= 2023) %>%
  group_by(MaterialType) %>%
  summarize(TotalCheckouts = sum(Checkouts))

# Create pie chart
ggplot(material_counts, aes(x = "", y = TotalCheckouts, fill = MaterialType)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Material Type Breakdown of Checkouts (2019-2023)")

```